Xor-Some Preregistered Study: Analyses

### 2022-04-17 analysis/main/02_xor-some_prereg_plots_final.Rmd
create dist and posterior plots on simulated data

Below, some simulations are conducted to create plots for the paper.

>>>>

Below, some simulations are conducted to create plots for the paper. Simulate data similarly to the power analysis, assuming an idealized effect size of 0.2 (somewhat larger than in power analysis, for better visibility). Modify the sampling a little to create low and high condition data.
```{r}
# simulate data from 200 participants with the function from the power analysis
create_fake_data <- function(N = 100, seed = 1) {
  set.seed(seed)
  # optionally changed to 0.25
  beta_rel  <- 0.25
  beta_comp <- 0.25
  beta_pri  <- -0.25
  map_df(
    1:N, 
    function(i) {
      tibble(
        subj    = rep(str_c("subj_",i), times = 8),
        trigger = rep(c('or', 'some'), times= 4), 
        rel     = rnorm(8),
        comp    = rnorm(8),
        pri     = rnorm(8)
      ) %>% 
        group_by(subj) %>% 
        mutate(
          rel     = (rel-mean(rel)) / sd(rel),
          comp    = (comp-mean(comp)) / sd(comp),
          pri     = (pri-mean(pri)) / sd(pri)
        ) %>% 
        ungroup() %>% 
        mutate(
          target  = rnorm(8, beta_rel * rel + beta_comp * comp + beta_pri * pri)
        )
    }
  )
}

fake_data_scatter <- create_fake_data(200, 1234)
```

```{r}
fake_data_long <- fake_data_scatter %>% pivot_longer(
  c(comp, rel, pri), 
               names_to = "Predictors", 
               values_to = "Predictor_response"
) %>%
  mutate(
    prior_categorization = ifelse(Predictor_response <= 0, 0, 1),
    prior_categorization = as.factor(prior_categorization)
  ) %>%
    mutate(
      `Prior categorization` = factor(prior_categorization, levels = c(0,1),
                                         labels=c("low", "high")),
         Trigger = factor(trigger, levels = c("some", "or"), 
                            labels = c("Some", "Or")),
         Predictors = factor(Predictors, levels = c("rel", "comp", "pri"),
                             labels = c("Relevance", "Competence", "Prior"))
    )

# write out for fitting the idealized model in the analysis script
#fake_data_long %>% write_csv("../../data/main/idealized_simulation_long_N200_B025.csv")

fake_data_long %>%
  ggplot(., aes(x = Predictor_response, y = target)) +
  geom_point(aes(color = `Prior categorization`), size = 2, alpha = 0.8) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_grid(cols=vars(Predictors), rows=vars(Trigger)) +
  theme_bw() +
  labs(x = "Participants' factor ratings", y = "Participants' inference likelihood ratings", color = "Prior categorization of vignettes") +
  theme(legend.position = "top",
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 13),
        axis.title = element_text(size = 13),
        strip.text.x = element_text(size = 13),
        strip.text.y = element_text(size = 13)) 
#ggsave("../../writing/paper/images/idealized_data_N200_b025.pdf",  width = 9, height = 7)
```

### 2022-04-17 analysis/main/03_xor-some_prereg_analysis_final.Rmd
create dist and posterior plots on simulated data

>>>>

### Analysis on fake simulated data for visualization purposes
No random effects here since none were simulated.
```{r}
fake_data <- read_csv("../../data/main/idealized_simulation_long_N200_B025.csv")

# Full model 
fake_data_wide <- fake_data %>% select(-prior_categorization, -`Prior categorization`, -Trigger ) %>% pivot_wider(
  names_from = Predictors, values_from = Predictor_response
) %>% 
  mutate(
    pri = Prior,
    comp = Competence,
    rel = Relevance,
  main_type = factor(trigger, levels=c("some", "or"))
)

# some = 0, xor = 1
contrasts(fake_data_wide$main_type)

model_fake <- brm(target ~ pri*comp*rel* main_type,
                 data = fake_data_wide,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.9),
                 cores = 4,
                 iter = 3000)

# posteriors
model_fake %>% spread_draws(b_Intercept, b_pri, b_comp, b_rel,
                          b_main_typeor, `b_pri:comp`,
                          `b_pri:rel`, `b_comp:rel`,
                          `b_pri:main_typeor`, `b_comp:main_typeor`,
                          `b_rel:main_typeor`, `b_pri:comp:rel`, 
                          `b_pri:comp:main_typeor`, `b_pri:rel:main_typeor`,
                          `b_comp:rel:main_typeor`, 
                          `b_pri:comp:rel:main_typeor`) %>% 
  mutate(
    prior_xor = b_pri+ `b_pri:main_typeor`,
    prior_some = b_pri,
    competence_xor = b_comp + `b_comp:main_typeor`,
    competence_some = b_comp,
    relevance_xor =  b_rel + `b_rel:main_typeor`,
    relevance_some = b_rel
  ) -> model_fake_posteriors

```{r}
model_fake_posteriors %>% select(prior_xor, prior_some,
                               competence_xor, competence_some,
                               relevance_xor, relevance_some) %>%
  pivot_longer(cols = everything(), names_to = "effect", values_to = "value") -> model_fake_posteriors_long

model_fake_posteriors_long %>%
  mutate(effect = factor(effect, 
                         levels = c("relevance_xor", "relevance_some", "prior_xor", "prior_some", "competence_xor", "competence_some"), 
                         labels = c("Relevance (or)", "Relevance (some)", "Prior (or)", "Prior (some)", "Competence (or)", "Competence (some)"))) %>%
  ggplot(.) + # 
  stat_halfeye(aes(y = effect, x = value, 
                   fill = stat(case_when(x %>% between(-0.05, 0.05) ~ "no effect", x < -0.05 ~ "negative effect", x > 0.05 ~ "positive effect"))
                   ), 
               alpha = .7) +
  geom_vline(xintercept = c(-.05, .05), linetype = "dashed") +
  scale_fill_manual(name = "Direction of effect", values = c("coral", "mediumseagreen")) + # , "yellow3"
  labs(x = "Effect size", y = "Effect") +
  theme(legend.position = "top",
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        axis.title = element_text(size = 13),
        strip.text.x = element_text(size = 13),
        strip.text.y = element_text(size = 13))

#ggsave("../../writing/paper/images/posterior-effects-idealized-sim.pdf", width=7, height=5)
```
